Factoring with $n+2$ clean qubits and $n-1$ dirty qubits

### 2017-06-23 doc/paper.tex
Addressing Marissa's minor comment and Austin's major comment.

\begin{abstract}
We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae in an unknown state that must be preserved.
We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
\end{abstract}

@@

In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
We do so without increasing the asymptotic costs, though the constant factors of our constructions are worse.

@@

Our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
Achieving small constant factors under realistic architectural constraints is an important practical problem, but not the focus of this paper.

@@

Possible future improvements aside, in this paper we showed that a number of clean qubits sufficient to perform Shor's algorithm is $n + O(1)$.
We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
Such a construction could, by allowing factoring problems to fit into quantum computers with half as much space, reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly quantum computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).

@@

We thank Matthew Neeley and Dave Bacon for comments on earlier versions of this paper, from which it greatly benefited.

>>>>

\begin{abstract}
We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e. qubits in an unknown state that must be preserved).
We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
\end{abstract}

@@

In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
We do so without increasing the asymptotic costs.
However, our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
Achieving low T gate counts under realistic architectural constraints is an important practical problem, but not the focus of this paper.

@@

@@

Possible future improvements aside, in this paper we showed that the number of clean qubits sufficient to perform Shor's algorithm is no more than $n + O(1)$.
We consider this to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
We do note, however, that these improvements are of more theoretical than practical interest.
In practice, on error corrected quantum computers, the important cost to optimize is the number of T gates.
In that sense, the main {\em practical} contribution of this paper is the improved increment circuit.

@@

We thank Matthew Neeley, Dave Bacon, and Austin Fowler for comments on earlier versions of this paper, from which it greatly benefited.
